Implementing Real-Time Transport Services over an Ossified Network

### 2016-05-23 paper.tex
switch to 10pt + sig-alternet edit for compability with microtype

\documentclass{sig-alternate-05-2015}
\usepackage[l2tabu,orthodox]{nag}
\usepackage[utf8x]{inputenc}       
\usepackage[british]{babel}        
\usepackage[babel=true]{microtype}
\usepackage{amsmath}
\usepackage[all]{onlyamsmath}
\usepackage{newtxtext}
\usepackage{newtxmath}
\usepackage{upquote}
\usepackage{graphicx}
\usepackage{url}
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{bytefield}
\usepackage{listings}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{color}
\usepackage{balance}
\usepackage{tabularx}

@@

In the IETF, the Transport Services (TAPS) working group is chartered to 
(1) develop a taxonomy of \emph{transport services}, that is, to identify the
features that comprise, and can be combined, to form complete transport
protocols; and (2) to develop an abstract API for applications to request
desirable services, allowing the system to select an appropriate transport
protocol based on application needs. It is hoped that this will loosen the
coupling between application and transport, so enabling deployment of new 
transport protocols.

@@

The presence of deadlines, in a network that offers only a best-effort
packet deliver service, implies that the transport must provide
\emph{partial reliability}.
The network has non-zero probability of losing any particular packet. 
If forward error correction is used to repair loss, this implies there is
always some probability that a particular packet will be non-recoverable.
If retransmission is used to recover from losses, there is the potential 
for unbounded delay, since any retransmission may itself be lost. 
These probabilities can be estimated, to give probabilistic constraints
on loss and delay, but will be non-zero. Accordingly, to meet deadlines,
the transport must support partial reliability service, accepting that
it may not be able to deliver all data by its deadline.
Many real-time applications run over TCP today, and TCP does not offer a
partial reliability service. This results in occasional play-out stalls, 
when the transport blocks the application because a retransmission takes
too long. These stalls are one of the primary causes of poor user
experience in streaming applications, and are directly caused by the lack
of partial reliability in TCP. Missing one frame that has not been
delivered by its deadline is much less disruptive than a stall in play-out.

Given deadlines and partial reliability, the next important transport
service is \emph{dependency management}. There is no point sending data
that cannot be used, because it depends on previous data that was lost. 
\todo{...complex trade-off, since you might want to send data that misses its 
   deadline, if it allows later data to be used.}

\todo{...following from this, we have the concept of application-level framing
   to make the best use of data, irrespective of losses. This implies that
   the transport provides a \emph{message oriented} service that respects
   ADU boundaries.}

\todo{...we note that many multimedia applications make use of multiple streams. 
   For example, audio and video are logically separate, etc.
   can be supported by opening multiple transport connections, but
   desirable to support a \emph{sub-stream} service to reduce overheads.}

\todo{...finally, we note that \emph{congestion control} is important. 
   motivate: historically there has been a belief that real-time doesn't 
   do congestion control, needs an isochronous channel. 
   not practical for the Internet. 
   Also, experience has shown not necessary for many real-time applications.
   Yes, there are some that need constant bandwidth and cannot adapt, but
   many are adaptive. 
   Better user experience to adapt to fit the channel, rather than recover
   from loss after the channel discards packets that don't fit. }

@@

\begin{table*}
  \centering
  \begin{tabularx}{\textwidth}{llll}
    \toprule
      Transport Service            & Function & Parameters                                     & Return Value \\
    \midrule
                                   & \texttt{socket}         
                                              & \texttt{af} -- Address family                  & Socket descriptor \\
                                   &          & \texttt{st} -- Socket type & \\
                                   & \texttt{close}          
                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
    \midrule
      \textbf{Connection oriented} & \texttt{bind}           
                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
                                   &          & \texttt{addr} -- Address to bind to & \\
                                   &          & \texttt{addrlen} -- Length of \texttt{addr} & \\
                                   & \texttt{listen}         
                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
                                   & \texttt{accept}        
                                              & \texttt{sd} -- Listening socket descriptor     & Connection socket descriptor \\
                                   &          & \texttt{addr} -- Address of peer & \\
                                   &          & \texttt{addrlen} -- Length of \texttt{addr} & \\
                                   & \texttt{connect}        
                                              & \texttt{addr} -- Address to connect to         & 0 (success), -1 (error) \\
                                   &          & \texttt{addrlen} -- Length of \texttt{addr} & \\
    \midrule
      \textbf{Deadlines}           & \texttt{set\_po\_delay} 
                                              & \texttt{delay} -- Playout delay (in ms)        & 0 (success), -1 (error) \\
    \midrule
      \textbf{Message oriented}    & \texttt{send\_message}  
                                              & \texttt{sd} -- Socket descriptor               & Number of bytes sent \\
                                   &          & \texttt{buf} -- Message data & \\
                                   &          & \texttt{len} -- Length of message data & \\
                                   &          & \texttt{seq\_num} -- Sequence number & \\
      \emph{Deadlines}             &          & \texttt{deadline} -- Relative deadline of message (in ms) & \\ 
      \emph{Dependencies}          &          & \texttt{depends\_on} -- \texttt{seq\_num} of dependency & \\
      \emph{Sub-streams}           &          & \texttt{substream} -- Substream identifier & \\
                                   & \texttt{recv\_message}  
                                              & \texttt{sd} -- Socket descriptor               & Number of bytes received \\
                                   &          & \texttt{buf} -- Buffer for message data & \\
                                   &          & \texttt{len} -- Size of \texttt{buf} & \\
      \emph{Sub-streams}           &          & \texttt{substream} -- Substream identifier & \\
                                   &          & \todo{return value, rather than param?} & \\
    \bottomrule
  \end{tabularx}
  \caption{Outline transport API for real-time applications. Return values
  shown are for successful calls; in all cases, -1 is returned in the event
  of an error}
  \label{tab:api}
\end{table*}

@@

Evidence that these services can be deployed above UDP exists in the form
of the WebRTC data channel \cite{draft-ietf-rtcweb-data-channel-13}.  
\todo{expand}

\todo{also QUIC}

@@

Deployments using UDP are popular, and work well. However, as described in
Section \ref{sec:ossification}, there are also reasons for providing these 
services running over TCP. 
\todo{
  these services are deployed over TCP now, but have an inconvenient API that
  imposes lots of work on application developers, and with higher than desired
  latency. 
  TCP Hollywood offers a way to fix both issues.
  Brief summary.
  Obviously we need to change TCP congestion control too.
}

The architecture of TCP Hollywood \cite{mcquistin2016hollywood} is shown in
Figure \ref{diagram:tcp-hollywood}. TCP Hollywood provides all of the
services described in Section \ref{sec:services}, with functionality split
across an intermediary layer (in userspace), and a set of modifications to
the kernel. The intermediary layer is responsible for providing the
message-oriented abstraction, and exposing the API for partial reliability.
The kernel modifications implement the deadline and dependency services,
as well as enabling out-of-order delivery of incoming segments. This split
allows applications to use a single API, whether or not the kernel
modifications are available. TCP Hollywood inherits standard TCP's
congestion control mechanism. Clearly, this loss-based mechanism is not
suitable for real-time applications, and a delay-based algorithm should
be adopted.

@@

@@

%==================================================================================================
\section{Acknowledgements}

>>>>

\documentclass[10pt]{sig-alternate-05-2015}
\usepackage[l2tabu,orthodox]{nag}
\usepackage[utf8x]{inputenc}
\usepackage[british]{babel}
\usepackage[babel=true]{microtype}
\usepackage{amsmath}
\usepackage[all]{onlyamsmath}
\usepackage{newtxtext}
\usepackage{newtxmath}
\usepackage{upquote}
\usepackage{graphicx}
\usepackage{url}
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{bytefield}
\usepackage{listings}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{color}
\usepackage{balance}
\usepackage{tabularx}

@@

In the IETF, the Transport Services (TAPS) working group is chartered to
(1) develop a taxonomy of \emph{transport services}, that is, to identify the
features that comprise, and can be combined to form, complete transport
protocols; and (2) to develop an abstract API for applications to request
desirable services, allowing the system to select an appropriate transport
protocol based on application needs. It is hoped that this will loosen the
coupling between application and transport, so enabling deployment of new
transport protocols.

@@

The presence of deadlines, in a network that offers only a best-effort
packet delivery service, implies that the transport must provide
\emph{partial reliability}.
The network has non-zero probability of losing any particular packet.
If forward error correction is used to repair loss, this implies there is
always some probability that a particular packet will be non-recoverable.
If retransmission is used to recover from losses, there is the potential
for unbounded delay, since any retransmission may itself be lost.
These probabilities can be estimated, to give probabilistic constraints
on loss and delay, but will be non-zero. Accordingly, to meet deadlines,
the transport must support partial reliability service, accepting that
it may not be able to deliver all data by its deadline.
Many real-time applications run over TCP today, and TCP does not offer a
partial reliability service. This results in occasional play-out stalls,
when the transport blocks the application because a retransmission takes
too long. These stalls are one of the primary causes of poor user
experience in streaming applications, and are directly caused by the lack
of partial reliability in TCP. Missing one frame that has not been
delivered by its deadline is much less disruptive than a stall in play-out.

Given deadlines and partial reliability, the next important transport
service is \emph{dependency management}. There is no point in sending data
that cannot be used because it depends on previous data that was not
received. Providing this service is complicated by the two ways in which
data can be \emph{useful} to applications: it may itself by played out, or
it may be needed as part of the application's decoding chain. For example,
an I frame provides utility both in being played out, and in allowing for
subsequent P and B frames. As a result, dependency information may result
in deadline information being overruled: data may be sent even if it cannot
be played out, if it is necessary for the decoding of other data.

Providing partial reliability, through deadlines and dependencies, requires
the use of application-level framing \cite{clark:1990:architecture} to
make the best use of payload data, irrespective of packet loss. At the
transport layer, this implies that we provide a \emph{message oriented}
service, that maintains ADU boundaries.

Many multimedia applications make use of multiple data flows. For example,
a simple IPTV application will maintain separate audio and video flows.
These could be sent across multiple transport-layer connections, but
overheads can be reduced by multiplexing these flows on a single
connection. A \emph{sub-stream} service allows this.

We note that \emph{congestion control} is important. Clearly, the historic
belief that real-time applications require an isochronous channel, and
cannot be subjected to congestion control, is not practical on the
Internet. Further, while there is a number of applications that are not
adaptive (and require constant bandwidth), an increasing number are
adaptive, and would better serve users by adapting to the available
bandwidth.

@@

\begin{table*}
  \centering
  \begin{tabularx}{\textwidth}{llll}
    \toprule
      Transport Service            & Function & Parameters                                     & Return Value \\
    \midrule
                                   & \texttt{socket}
                                              & \texttt{af} -- Address family                  & Socket descriptor \\
                                   &          & \texttt{st} -- Socket type & \\
                                   & \texttt{close}
                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
    \midrule
      Connection oriented          & \texttt{bind}
                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
                                   &          & \texttt{addr} -- Address to bind to & \\
                                   &          & \texttt{addrlen} -- Length of \texttt{addr} & \\
                                   & \texttt{listen}
                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
                                   & \texttt{accept}
                                              & \texttt{sd} -- Listening socket descriptor     & Connection socket descriptor \\
                                   &          & \texttt{addr} -- Address of peer & \\
                                   &          & \texttt{addrlen} -- Length of \texttt{addr} & \\
                                   & \texttt{connect}
                                              & \texttt{addr} -- Address to connect to         & 0 (success), -1 (error) \\
                                   &          & \texttt{addrlen} -- Length of \texttt{addr} & \\
    \midrule
      Deadlines                    & \texttt{set\_po\_delay}
                                              & \texttt{delay} -- Playout delay (in ms)        & 0 (success), -1 (error) \\
    \midrule
      Message oriented             & \texttt{send\_message}
                                              & \texttt{sd} -- Socket descriptor               & Number of bytes sent \\
                                   &          & \texttt{buf} -- Message data & \\
                                   &          & \texttt{len} -- Length of message data & \\
                                   &          & \texttt{seq\_num} -- Sequence number & \\
      Deadlines                    &          & \texttt{deadline} -- Relative deadline of message (in ms) & \\
      Dependencies                 &          & \texttt{depends\_on} -- \texttt{seq\_num} of dependency & \\
      Sub-streams                  &          & \texttt{substream} -- Substream identifier & \\
                                   & \texttt{recv\_message}
                                              & \texttt{sd} -- Socket descriptor               & Number of bytes received \\
                                   &          & \texttt{buf} -- Buffer for message data & \\
                                   &          & \texttt{len} -- Size of \texttt{buf} & \\
      Sub-streams                  &          & \texttt{substream} -- Substream identifier & \\
                                   &          & \todo{return value, rather than param?} & \\
    \bottomrule
  \end{tabularx}
  \caption{Outline transport API for real-time applications. Return values
  shown are for successful calls; in all cases, -1 is returned in the event
  of an error}
  \label{tab:api}
\end{table*}

@@

Evidence that these services can be deployed above UDP exists in the form
of the WebRTC data channel \cite{draft-ietf-rtcweb-data-channel-13} and
QUIC protocol \cite{draft-tsvwg-quic-protocol-02}.
The former is a peer-to-peer protocol, comprising an SCTP association
running over DTLS, itself running over a UDP flow negotiated via an SDP
\cite{RFC4566} offer/answer exchange \cite{RFC3264} as part of a WebRTC
session \cite{jennings:2013:rtcweb} (the WebRTC media channel uses RTP
running over UDP also, further showing the utility of UDP-based data).
This has been deployed in popular web browsers, with global deployment, and
demonstrated to be effective.  The latter is implemented by Google in their
Chrome browser, and used as an alternative to TCP has a significant
fraction of web traffic downloads from their domain.

@@

Deployments using UDP are popular, and work well. However, as described in
Section \ref{sec:ossification}, there are also reasons for providing these
services over TCP, since there are a significant fraction of networks that
block UDP traffic. It is clearly possible to run real-time traffic over
TCP, as demonstrated by applications such as NetFlix or the BBC iPlayer
that comprise the majority of Internet traffic. However, TCP has a
inconvenient API that imposes lots of work on application developers, and
introduces higher than desired latency. We have shown how to address these
issues, and provide the full set of transport services we propose in
Section \ref{sec:services} in previous work, with our TCP Hollywood
proposal \cite{mcquistin2016hollywood}.

@@

The architecture of TCP Hollywood is shown in
Figure \ref{diagram:tcp-hollywood}. TCP Hollywood provides all of the
services described in Section \ref{sec:services}, with functionality split
across an intermediary layer (in user-space), and a set of modifications to
the kernel. The intermediary layer is responsible for providing the
message-oriented abstraction, and exposing the API for partial reliability.
The kernel modifications implement the deadline and dependency services,
as well as enabling out-of-order delivery of incoming segments. This split
allows applications to use a single API, whether or not the kernel
modifications are available. TCP Hollywood inherits standard TCP's
congestion control mechanism. Clearly, this loss-based mechanism is not
suitable for real-time applications, and a delay-based algorithm should
be adopted.

\todo{Stephen: give more detail, and expand this to a whole column}

\todo{discuss how we need to change TCP congestion control}

@@

%==================================================================================================
%\section{Acknowledgements}
